
## Getting and Cleaning Data Course Project Codebook

### Importing relevant data sets

After successfully setting up the working directory (using `setwd()`), the .txt files were imported to R through the `read.table` function. 

```{r importing}
test_s <- read.table("test/subject_test.txt")
test_X <- read.table("test/X_test.txt")
test_Y <- read.table("test/Y_test.txt")
train_s <- read.table("train/subject_train.txt")
train_X <- read.table("train/X_train.txt")
train_Y <- read.table("train/Y_train.txt")
features <- read.table("features.txt")
```

### Renaming variable names for test and training sets

Note that the `features` data set contains the variable names for the measurements. The data frame `features` is then attached to the test and training data set for the measurements (`test_X` and `train_X`, respectively).

```{r naming}
colnames(test_X) <- features[,2]
colnames(train_X) <- features[,2]
```

The six activities (walking, walking upstairs, etc.) are still encoded in numerical values, thus a function was created (called `make_label`) which replaces the numbers with the actual names of the activities provided. This function was used to create labels for the test and train data set activities (which are stored in the `test_Y` and `train_Y` data sets, respectively).

```{r labeling}
make_label <- function(df){
    label <- rep(0,nrow(df))
    for(i in 1:nrow(df)){
        label[i] <- switch(df[i,],"WALKING","WALKING_UPSTAIRS","WALKING_DOWNSTAIRS",
                           "SITTING","STANDING","LAYING")
    }
    return(label)
}
test_labels <- make_label(test_Y)
train_labels <- make_label(train_Y)
```

### Merging data sets and some further renaming

Now, the .txt files that corresponds to the test and train data sets are then combined. Note that the first three columns of both the test and train data frames contain the respondent tag, the activity performed, and whether they belong to the test set or training set, respectively. Then finally the test set and training set are merged together using `rbind` to a data set called `mydata`.

```{r}
test <- data.frame(test_s,test_labels,rep("test",nrow(test_X)),test_X)
train <- data.frame(train_s, train_labels,rep("train",nrow(train_X)),train_X)
colnames(test)[1:3] <- c("respondent","activity","group")
colnames(train)[1:3] <- c("respondent","activity","group")
mydata <- rbind(test, train)
```

Relating the concept of tidy data for this data set, a long data set was created, where the first three variables are the descriptive variables with the following variables as the measurements. The categories of these variables are not broken down to separate variables (which creates a wide data set).

### Summarizing the combined test and training data sets using mean and standard deviation

Finally, the mean and standard errors of each measurements for all activities of each of the respondents were obtained, using functions from the `dplyr` package.

```{r message = F}
library(dplyr)
mymeans <- mydata %>% group_by(respondent, activity, group) %>% summarise_all(funs(mean))
mysd <- mydata %>% group_by(respondent, activity, group) %>% summarise_all(funs(sd))
```
 
